---
title: "Anna_primary_bclass"
author: "Anna Liang"
date: "2025-11-16"
output: pdf_document
---

# Develop predictive model for binary classfication.
You can use any preprocessing and any modeling technique you like, including ones not discussed in class (consider exploring RNN or SVM). Export a deployable copy of each model and a set of predictions from each model on the test data claims-test.RData. Predictions should be formatted as a single data frame named pred_df with columns .id containing the URL ID, bclass.pred containing the predicted label for the binary class setting, mclass.pred containing the predicted label for the multiclass setting

Unigram + Bigram TF–IDF + Regularized Logistic Regression (Binary Model)



Load packages.
```{r}
library(tidyverse)
library(text2vec)
library(here)

source(here("scripts/preprocessing.R"))
set.seed(1234)
```

Load raw training data and clean
```{r}
load(here("data/claims-raw.RData")) # gives claims_raw

claims_clean_binary <- claims_raw %>%
  parse_data() %>%                     # adds text_clean column
  select(.id, text_clean, bclass) %>%  
  mutate(bclass = factor(bclass),
         label_binary = ifelse(bclass == "Relevant claim content", 1L, 0L)) 

# 1 = claims, 0 = everything else
# columns: .id, text_clean, bclass, label_binary

# number of rows
nrow(claims_clean_binary)

save(claims_clean_binary, file = here("data/claims-clean-binary.RData"))
```

Create TF–IDF features for binary training
```{r}
# Load cleaned training data
load(here("data/claims-clean-binary.RData"))

# Create text iterator
it_train <- itoken(
  claims_clean_binary$text_clean,
  ids = claims_clean_binary$.id,
  progressbar = FALSE
)
```

Build vocabulary (unigrams + bigrams)
```{r}
vocab_bi <- create_vocabulary(
  it_train,
  ngram = c(1L, 2L)    # unigrams + bigrams
)

# Prune rare terms
vocab_bi <- prune_vocabulary(
  vocab_bi,
  term_count_min = 5L # Remove terms that appear less than 5 times
)

# Vocabulary size
nrow(vocab_bi) 
```

Create TF–IDF matrix
```{r}
vectorizer_bi <- vocab_vectorizer(vocab_bi)

# document-term matrix (DTM)
dtm_train_bi <- create_dtm(it_train, vectorizer_bi)

# rows = documents (webpages)
# columns = vocabulary terms (unigrams & bigrams)
# entries = raw term counts per document

# TF-IDF
tfidf_bi <- TfIdf$new()

X_train_bi <- tfidf_bi$fit_transform(dtm_train_bi) # Feature matrix
y_train    <- claims_clean_binary$bclass

# TF-IDF matrix dimensions
cat("TF-IDF matrix:", dim(X_train_bi)[1], "documents", ",", dim(X_train_bi)[2], "terms\n")
```


Attempt 1: Fitting a SVM classifier w/ linear kernel was not possible due to the amount of features we have.
Even after pruning to reduce the maximum number of vocabulary to 10000, running the model made rstudio crash.
```{r}
#vocab_bi <- prune_vocabulary(
#  vocab_bi,
#  term_count_min = 5L, # Remove terms that appear less than 5 times
#  doc_proportion_max = 0.3, # Drop terms that appear in >30% of documents
#  vocab_term_max = 10000L 
# )

#svm_bi <- svm(
#  x = as.matrix(X_train_bi),
#  y = claims_clean_binary$bclass, # factor labels
#  kernel = "linear",
#  probability = FALSE,
#  scale = FALSE
# )
```

Attempt 2: Regularized Logistic Regression (Glmnet) model

```{r}
library(glmnet)

library(Matrix)

X_glm <- as(X_train_bi, "dgCMatrix")
y_train_bi <- as.numeric(claims_clean_binary$label_binary)

# Cross validation
bin_glmnet <- cv.glmnet(
  x            = X_glm,
  y            = y_train_bi,
  family       = "binomial",
  type.measure = "class",
  nfolds       = 5
)


pred_glmnet <- predict(
  bin_glmnet,
  X_glm,
  s    = "lambda.min",
  type = "class"
)[, 1]

acc_glmnet <- mean(pred_glmnet == y_train_bi)
cat("Training accuracy (glmnet):", round(acc_glmnet, 3), "\n")

```


```{r}
# Save glmnet model
model_bundle <- list(
  glmnet_model = bin_glmnet,
  vocabulary   = vocab_bi,
  vectorizer   = vectorizer_bi,
  tfidf_model  = tfidf_bi
)

# Save model and vectorizers
save(model_bundle, file = here("results/bclass_primary_model.RData"))
```

#Applying this model to the test set

```{r}
# Load trained model
load(here("results/bclass_primary_model.RData"))
```


Load test set and preprocess
```{r}
load(here("data/claims-test.RData"))

claims_test_clean <- claims_test %>%
  parse_data() %>%           
  select(.id, text_clean)

```

Build TF–IDF for test set
```{r}
it_test <- itoken(
  claims_test_clean$text_clean,
  ids         = claims_test_clean$.id,
  progressbar = FALSE
)

# Use the trained vectorizer (fixed vocabulary)
dtm_test <- create_dtm(it_test, model_bundle$vectorizer)

# Apply the trained TF–IDF transformer (same idf as training)
X_test <- model_bundle$tfidf_model$transform(dtm_test)

```

Predict binary labels on test set
```{r}
# cv.glmnet: use s = "lambda.min"
p_hat_test <- predict(
  model_bundle$glmnet_model,
  newx = X_test,
  s    = "lambda.min",
  type = "response"
)[, 1]

# 0/1 predictions with 0.5 cutoff
pred_bin_test <- ifelse(p_hat_test > 0.5, 1L, 0L)

# Map 0/1 → actual label strings
bclass_pred <- ifelse(
  pred_bin_test == 1L,
  "Relevant claim content",
  "N/A: No relevant content."
)

```

Build prediction data frame
```{r}
pred_df_binary <- tibble(
  .id         = claims_test_clean$.id,
  bclass.pred = bclass_pred
)
```

```{r}
# Saving the binary predictions
if (!dir.exists(here("results"))) dir.create(here("results"), recursive = TRUE)
save(pred_df_binary, file = here("results/binaryclass_results/predictions.RData"))

head(pred_df_binary)
```


Sensitivity, Specificity, and Accuracy.
```{r}
# Make sure both are plain 0/1 numeric
truth <- as.numeric(y_train_bi)          # should already be 0/1
pred  <- as.numeric(pred_glmnet)         # should be 0/1 from glmnet

# Confusion matrix (0 = N/A, 1 = Relevant)
cm <- table(
  truth = factor(truth, levels = c(0, 1)),
  pred  = factor(pred,  levels = c(0, 1))
)
cm

# Extract cells
TN <- cm["0", "0"]  # true negatives  (N/A predicted N/A)
FP <- cm["0", "1"]  # false positives (N/A predicted Relevant)
FN <- cm["1", "0"]  # false negatives (Relevant predicted N/A)
TP <- cm["1", "1"]  # true positives  (Relevant predicted Relevant)

# Metrics
accuracy_cm  <- (TP + TN) / sum(cm)
sensitivity  <- TP / (TP + FN)   # recall, TPR for "Relevant claim content"
specificity  <- TN / (TN + FP)   # TNR for "N/A: No relevant content."

cat("Accuracy from confusion matrix:", round(accuracy_cm, 3), "\n")
cat("Sensitivity (TPR):", round(sensitivity, 3), "\n")
cat("Specificity (TNR):", round(specificity, 3), "\n")

cat("Accuracy from mean():", round(acc_glmnet, 3), "\n")
cat("Accuracy from CM   :", round(accuracy_cm, 3), "\n")
```

